R Notebook

### 2016-10-03 manuscript/draft.Rmd

Quality of predictions is important:
  * helps determine policy/set priorities
  * something something climate change
  * need to know both accuracy of point estimates and how reasonable our prediction intervals are.
  * Overconfident predictions can be worse than uncertainty
  * Quality of prediction is one metric for how much we know about ecological processes
Dominant method is stacked SDMs:
  *
Difficult to test model assumptions/predictions because of time scales
  * Most scientists making predictions today will have retired by 2050
  * Use long-term data sets such as BBS to evaluate predictive skill
    * Note the large change in average temperatures since the 80's; comparable to the changes expected in the next N years
  * Test with simple baselines (e.g. random walk, future-is-like-present)
  * Focus on richness/diversity because \[citations saying that it's important\]

@@

* Mixed model for observer effort \& site effects
* Naive AR1 model (adjusted for observer effects)
* species-level GBM w/ mixed model point estimates as bonus predictors where available
* richness-level GBM w/ mixed model point estimates as bonus predictors where available

(need a side note about why we're not using a proper count distribution; underdispersion)

@@

* Temporal richness variability doesn't track the predictors we use; sites have "inertia"
* More important to use a reasonable error model than to get the best predictors \& basis expansions
* We've posted predictions for the next N years of BBS data, using several methods, and invite others to do so as well.
* In many ways, this is the best-case scenario for SDMs:
  * large data set, known \& consistent sampling technique over a period of decades
  * True absences
  * extended predictor set
  * birds arguably migrate quickly enough to make equlibrium \& especially space-for-time assumptions less ridiculous
  * errors for individual species can cancel out; only asking for aggregate values
* Value of stacking versus predicting the quantitity of interest directly

>>>>

Quality of predictions is important:
  * helps determine policy/set priorities
  * something something climate change
  * need to know both accuracy of point estimates and how reasonable our prediction intervals are.
  * Overconfident predictions can be worse than uncertainty
  * Quality of prediction is one metric for how much we know about ecological processes
Unclear how much to trust existing predictions:
  * Strong assumptions (see "misc" below)
  * many *interesting* predictions
  * Potential problems with confidence intervals
  * Potential disconnect between training target and actual goals
  * Difficult to test model assumptions/predictions because of time scales
    * Most scientists making predictions today will have retired by 2050
Demonstrate best practices \& their contribution to better predictions
  * Use long-term data sets such as BBS to evaluate predictive skill
  * Note the large change in average temperatures since the 80's; comparable to the changes expected in the next N years
  * Lots of talk about how we should do it, fewer examples of doing them all together
  * Examples of best practices
    * not just bioclim
    * time series
    * spatial/site-level stuff
    * probabilistic/error bars
    * multiple models
    * Test with simple baselines (e.g. random walk, future-is-like-present)
  * Focus on richness/diversity because \[citations saying that it's important\]
  * Publish forecasts at different timescales to evaluate performance/credibility
Here we do that:
  * forecasts for biodiversity in NAmerican birds with as many best practices as possible
  * Discuss the influence of these improved methods \& of our error rates on our ability to understand:
    * How to improve
    * How much to trust the predictions

@@

* Mixed model for observer effort \& site effects
* Naive AR1 model (adjusted for observer effects)
* species-level GBM w/ mixed model point estimates as bonus predictors where available
* richness-level GBM w/ mixed model point estimates as bonus predictors where available
* JSDM

(need a side note about why we're not using a proper count distribution; underdispersion, easier access to out-of-the-box time series methods)

@@

* Temporal richness variability doesn't track the predictors we use; sites have "inertia"
* More important to use a reasonable error model than to get the best predictors \& basis expansions
* We've posted predictions for the next N years of BBS data, using several methods, and invite others to do so as well.
* In many ways, this is the best-case scenario for SDMs:
  * large data set, known \& consistent sampling technique over a period of decades
  * True absences
  * extended predictor set
  * birds arguably migrate quickly enough to make equlibrium \& especially space-for-time assumptions less ridiculous
  * errors for individual species can cancel out; only asking for aggregate values
* None of the models actually work all that well, compared to baseline.
* Value of stacking versus predicting the quantitity of interest directly
* Maybe richness is too coarse
  * abundance
  * species-level
